---
title: "Protocol"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library

```{r library}
library(dplyr)
library(Seurat)
library(data.table)
# load data from Seurat package
#data(cc.genes)

# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

## Load the dataset

```{r load}
feature_counts.data <- read.table(file = "/Users/annaasklof/Desktop/Seurat/S/Data/feature_counts.table.tsv", 
                                  sep = "\t",
                                  header = TRUE,
                                  quote = "",
                                  stringsAsFactors = FALSE)
dim(x=feature_counts.data)
head(x = colnames(x = feature_counts.data))
#feature_counts.data
```

## Remove data that is not sc from dataset

```{r sc}
removeIDs <- function(feature_counts.data) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(feature_counts.data)) {
    if(grepl(".dil.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
    if(grepl(".FACS.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
      if(grepl(".bulk.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- feature_counts.data[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

feature_counts_SC <- removeIDs(feature_counts.data)
dim(feature_counts_SC)
head(x = colnames(x = feature_counts_SC))

```
## Save  dil data

```{r save dil}

# Dil
removeIDs <- function(feature_counts.data) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(feature_counts.data)) {
    if(grepl("sc", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
    if(grepl(".FACS.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
      if(grepl(".bulk.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- feature_counts.data[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

feature_counts_Dil <- removeIDs(feature_counts.data)
dim(feature_counts_Dil)
head(x = colnames(x = feature_counts_Dil))

```
## Save  bulk data

```{r save bulk}

# Bulk
removeIDs <- function(feature_counts.data) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(feature_counts.data)) {
    if(grepl("sc", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
    if(grepl(".FACS.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
      if(grepl(".dil.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- feature_counts.data[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

feature_counts_Bulk <- removeIDs(feature_counts.data)
dim(feature_counts_Bulk)
head(x = colnames(x = feature_counts_Bulk))

```

## Remove columns that are not of interest from dataset

```{r removeColumns}
# SC
feature_counts_SC= feature_counts_SC[,-2:-6]
dim(feature_counts_SC)
head(x = colnames(x = feature_counts_SC))
head(x = feature_counts_SC)

# Dil 
feature_counts_Dil= feature_counts_Dil[,-2:-6]
dim(feature_counts_Dil)
head(x = colnames(x = feature_counts_Dil))
head(x = feature_counts_Dil)

# Bulk 
feature_counts_Bulk= feature_counts_Bulk[,-2:-6]
dim(feature_counts_Bulk)
head(x = colnames(x = feature_counts_Bulk))
head(x = feature_counts_Bulk)

```

## Load file for correct gene names

```{r load2}
geneInfoFile <- read.table(file = "/Users/annaasklof/Desktop/Seurat/S/Data/geneInfo.QC.method.seurat.SCORPIUS.tsv", 
                           sep = "\t",
                           header = TRUE,
                           quote = "",
                           stringsAsFactors = FALSE)

#only protein coding genes
geneInfoFile = geneInfoFile[geneInfoFile$gene_biotype == "protein_coding",] 

#only genes represented with one unique ensembl ID

tmp = data.frame(table(geneInfoFile$external_gene_name))
tmp = tmp[tmp$Freq ==1, ]
geneInfoFile = geneInfoFile[geneInfoFile$external_gene_name %in% tmp$Var1, ]


dim(x=geneInfoFile)
```

## Select relevant columns

```{r select}

geneInfoNames = geneInfoFile %>%  select(Geneid, external_gene_name)
head(geneInfoNames)
dim(x=geneInfoNames)

intersect( g2m.genes, geneInfoNames$external_gene_name)
#intersect( s.genes, rownames(geneInfo))

```


## Perform inner join 

```{r innerJoin}

#Comb and Ref
geneInfo = inner_join(geneInfoNames, feature_counts_SC)
intersect( g2m.genes, geneInfo$external_gene_name)
dim(x=geneInfo)

# Dil
geneInfo_Dil = inner_join(geneInfoNames, feature_counts_Dil)
intersect( g2m.genes, geneInfo_Dil$external_gene_name)
dim(x=geneInfo_Dil)

# Bulk 
geneInfo_Bulk = inner_join(geneInfoNames, feature_counts_Bulk)
intersect( g2m.genes, geneInfo_Bulk$external_gene_name)
dim(x=geneInfo_Bulk)

```

## Set rownames to correct gene names

```{r rownames}
#test = data.frame(make.names(geneInfo$Chr, unique =TRUE)
#length(test)

#Comb and Ref
rownames(geneInfo) =  geneInfo$external_gene_name
dim(x=geneInfo)
head(x = colnames(x = geneInfo))

# Dil 
rownames(geneInfo_Dil) =  geneInfo_Dil$external_gene_name
dim(x=geneInfo_Dil)
head(x = colnames(x = geneInfo_Dil))

# Bulk 
rownames(geneInfo_Bulk) =  geneInfo_Bulk$external_gene_name
dim(x=geneInfo_Bulk)
head(x = colnames(x = geneInfo_Bulk))

```

## Remove columns in order to perform rowSums

```{r removeColumns2}

# Comb and Ref
geneInfo= geneInfo[,-1:-2]
dim(geneInfo)
head(x = colnames(x = geneInfo))
intersect( g2m.genes, rownames(geneInfo))

# Dil 
geneInfo_Dil= geneInfo_Dil[,-1:-2]
dim(geneInfo_Dil)
head(x = colnames(x = geneInfo_Dil))
intersect( g2m.genes, rownames(geneInfo_Dil))

# Bulk 
geneInfo_Bulk= geneInfo_Bulk[,-1:-2]
dim(geneInfo_Bulk)
head(x = colnames(x = geneInfo_Bulk))
intersect( g2m.genes, rownames(geneInfo_Bulk))

```

## Remove genes that are not expressed

```{r expressed}

#Comb and Ref
#rowSums(geneInfo)
keep <- rowSums(geneInfo)>1
geneInfo <- geneInfo[keep,]
dim(geneInfo)
intersect( g2m.genes, rownames(geneInfo))

# Dil 
keep_Dil <- rowSums(geneInfo_Dil)>1
geneInfo_Dil <- geneInfo_Dil[keep_Dil,]
dim(geneInfo_Dil)
intersect( g2m.genes, rownames(geneInfo_Dil))

# Bulk 
keep_Bulk <- rowSums(geneInfo_Bulk)>1
geneInfo_Bulk <- geneInfo_Bulk[keep,]
dim(geneInfo_Bulk)
intersect( g2m.genes, rownames(geneInfo_Bulk))

```

## Separate Comb and Ref

```{r Separate}
#Comb
removeIDs <- function(geneInfo) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(geneInfo)) {
    if(grepl("Ref", colnames(geneInfo[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- geneInfo[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

geneInfo_Comb <- removeIDs(geneInfo)
dim(geneInfo_Comb)
head(x = colnames(x = geneInfo_Comb))

#Ref
removeIDs <- function(geneInfo) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(geneInfo)) {
    if(grepl("Comb", colnames(geneInfo[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- geneInfo[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

geneInfo_Ref <- removeIDs(geneInfo)
dim(geneInfo_Ref)
head(x = colnames(x = geneInfo_Ref))

```
## Separate Comb_0h 

```{r Separate Comb_0h}
#Comb.0h
dim(geneInfo_Comb)
savedIDs <- function(geneInfo_Comb) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(geneInfo_Comb)) {
    if(grepl("Comb.0", colnames(geneInfo_Comb[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_savedColumns <- c()
  feature_counts_savedColumns <- geneInfo_Comb[,column_indices_ID]
  return(feature_counts_savedColumns)
}
geneInfo_Comb_0h<- savedIDs(geneInfo_Comb)
dim(geneInfo_Comb_0h)
head(x = colnames(x= geneInfo_Comb_0h))

```

## Separate Comb_24h 

```{r Separate Comb_24h}
#Comb.24h
dim(geneInfo_Comb)
savedIDs <- function(geneInfo_Comb) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(geneInfo_Comb)) {
    if(grepl("Comb.24", colnames(geneInfo_Comb[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_savedColumns <- c()
  feature_counts_savedColumns <- geneInfo_Comb[,column_indices_ID]
  return(feature_counts_savedColumns)
}
geneInfo_Comb_24h<- savedIDs(geneInfo_Comb)
dim(geneInfo_Comb_24h)
head(x = colnames(x= geneInfo_Comb_24h))

```

## Separate Comb_48h 

```{r Separate Comb_48h}
#Comb.48h
dim(geneInfo_Comb)
savedIDs <- function(geneInfo_Comb) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(geneInfo_Comb)) {
    if(grepl("Comb.48", colnames(geneInfo_Comb[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_savedColumns <- c()
  feature_counts_savedColumns <- geneInfo_Comb[,column_indices_ID]
  return(feature_counts_savedColumns)
}
geneInfo_Comb_48h<- savedIDs(geneInfo_Comb)
dim(geneInfo_Comb_48h)
head(x = colnames(x= geneInfo_Comb_48h))

```

## Create a Seurat object out of the table

```{r Seurat}

#Comb and Ref
#feature_counts_seurat <- CreateSeuratObject(counts = geneInfo, min.cells = 3)
#feature_counts_seurat
#feature_counts_seurat <- NormalizeData(object = feature_counts_seurat)
#feature_counts_seurat <- FindVariableFeatures(object = feature_counts_seurat, selection.method = "vst")
#feature_counts_seurat <- ScaleData(feature_counts_seurat, features = rownames(feature_counts_seurat))

#feature_counts_seurat <- RunPCA(object = feature_counts_seurat)
#DimHeatmap(feature_counts_seurat)

# Comb
feature_counts_seurat_Comb <- CreateSeuratObject(counts = geneInfo_Comb, min.cells = 3)
feature_counts_seurat_Comb
feature_counts_seurat_Comb <- NormalizeData(object = feature_counts_seurat_Comb)
feature_counts_seurat_Comb <- FindVariableFeatures(object = feature_counts_seurat_Comb, selection.method = "vst")
feature_counts_seurat_Comb <- ScaleData(feature_counts_seurat_Comb, features = rownames(feature_counts_seurat_Comb))
feature_counts_seurat_Comb <- RunPCA(object = feature_counts_seurat_Comb)
DimHeatmap(feature_counts_seurat_Comb)

# Comb 0h
feature_counts_seurat_Comb_0h <- CreateSeuratObject(counts = geneInfo_Comb_0h, min.cells = 3)
feature_counts_seurat_Comb_0h
feature_counts_seurat_Comb_0h <- NormalizeData(object = feature_counts_seurat_Comb_0h)
feature_counts_seurat_Comb_0h <- FindVariableFeatures(object = feature_counts_seurat_Comb_0h, selection.method = "vst")
feature_counts_seurat_Comb_0h <- ScaleData(feature_counts_seurat_Comb_0h, features = rownames(feature_counts_seurat_Comb_0h))
feature_counts_seurat_Comb_0h <- RunPCA(object = feature_counts_seurat_Comb_0h)
DimHeatmap(feature_counts_seurat_Comb_0h)

# Comb 24h
feature_counts_seurat_Comb_24h <- CreateSeuratObject(counts = geneInfo_Comb_24h, min.cells = 3)
feature_counts_seurat_Comb_24h
feature_counts_seurat_Comb_24h <- NormalizeData(object = feature_counts_seurat_Comb_24h)
feature_counts_seurat_Comb_24h <- FindVariableFeatures(object = feature_counts_seurat_Comb_24h, selection.method = "vst")
feature_counts_seurat_Comb_24h <- ScaleData(feature_counts_seurat_Comb_24h, features = rownames(feature_counts_seurat_Comb_24h))
feature_counts_seurat_Comb_24h <- RunPCA(object = feature_counts_seurat_Comb_24h)
DimHeatmap(feature_counts_seurat_Comb_24h)

# Comb 48h
feature_counts_seurat_Comb_48h <- CreateSeuratObject(counts = geneInfo_Comb_48h, min.cells = 3)
feature_counts_seurat_Comb_48h
feature_counts_seurat_Comb_48h <- NormalizeData(object = feature_counts_seurat_Comb_48h)
feature_counts_seurat_Comb_48h <- FindVariableFeatures(object = feature_counts_seurat_Comb_48h, selection.method = "vst")
feature_counts_seurat_Comb_48h <- ScaleData(feature_counts_seurat_Comb_48h, features = rownames(feature_counts_seurat_Comb_48h))
feature_counts_seurat_Comb_48h <- RunPCA(object = feature_counts_seurat_Comb_48h)
DimHeatmap(feature_counts_seurat_Comb_48h)

#Ref
feature_counts_seurat_Ref <- CreateSeuratObject(counts = geneInfo_Ref, min.cells = 3)
feature_counts_seurat_Ref
feature_counts_seurat_Ref <- NormalizeData(object = feature_counts_seurat_Ref)
feature_counts_seurat_Ref <- FindVariableFeatures(object = feature_counts_seurat_Ref, selection.method = "vst")
feature_counts_seurat_Ref <- ScaleData(feature_counts_seurat_Ref, features = rownames(feature_counts_seurat_Ref))
feature_counts_seurat_Ref <- RunPCA(object = feature_counts_seurat_Ref)
DimHeatmap(feature_counts_seurat_Ref)

# Dil 
feature_counts_seurat_Dil <- CreateSeuratObject(counts = geneInfo_Dil, min.cells = 3)
feature_counts_seurat_Dil
feature_counts_seurat_Dil <- NormalizeData(object = feature_counts_seurat_Dil)
feature_counts_seurat_Dil <- FindVariableFeatures(object = feature_counts_seurat_Dil, selection.method = "vst")
feature_counts_seurat_Dil <- ScaleData(feature_counts_seurat_Dil, features = rownames(feature_counts_seurat_Dil))
#feature_counts_seurat_Dil <- RunPCA(object = feature_counts_seurat_Dil)
#DimHeatmap(feature_counts_seurat_Dil)


# Bulk 
feature_counts_seurat_Bulk <- CreateSeuratObject(counts = geneInfo_Bulk, min.cells = 3)
feature_counts_seurat_Bulk
feature_counts_seurat_Bulk <- NormalizeData(object = feature_counts_seurat_Bulk)
feature_counts_seurat_Bulk <- FindVariableFeatures(object = feature_counts_seurat_Bulk, selection.method = "vst")
feature_counts_seurat_Bulk <- ScaleData(feature_counts_seurat_Bulk, features = rownames(feature_counts_seurat_Bulk))
#feature_counts_seurat_Bulk <- RunPCA(object = feature_counts_seurat_Bulk)
#DimHeatmap(feature_counts_seurat_Bulk)

```

## Assign cell-cycle scores

```{r Scoring}
s.genes

#Comb and Ref
#intersect( g2m.genes, rownames(geneInfo$external_gene_name))
#intersect( s.genes, rownames(geneInfo))

#feature_counts_seurat <- CellCycleScoring(feature_counts_seurat, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

#feature_counts_seurat@meta.data
#head(x = colnames(x = feature_counts_seurat))
#RidgePlot(feature_counts_seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
#feature_counts_seurat <- RunPCA(feature_counts_seurat, features = c(s.genes, g2m.genes))
#DimPlot(feature_counts_seurat)

# Comb
intersect( g2m.genes, rownames(geneInfo_Comb$external_gene_name))
intersect( s.genes, rownames(geneInfo_Comb))

feature_counts_seurat_Comb <- CellCycleScoring(feature_counts_seurat_Comb, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

#feature_counts_seurat_Comb@meta.data
#head(x = colnames(x = feature_counts_seurat_Comb))

feature_counts_seurat_Comb <- RunPCA(feature_counts_seurat_Comb, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat_Comb)

# Comb_0h 
intersect( g2m.genes, rownames(geneInfo_Comb_0h$external_gene_name))
intersect( s.genes, rownames(geneInfo_Comb_0h))

feature_counts_seurat_Comb_0h <- CellCycleScoring(feature_counts_seurat_Comb_0h, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Comb_0h$S.Score
feature_counts_seurat_Comb_0h$G2M.Score

feature_counts_seurat_Comb_0h <- RunPCA(feature_counts_seurat_Comb_0h, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat_Comb_0h)

# Comb_24h 
intersect( g2m.genes, rownames(geneInfo_Comb_24h$external_gene_name))
intersect( s.genes, rownames(geneInfo_Comb_24h))

feature_counts_seurat_Comb_24h <- CellCycleScoring(feature_counts_seurat_Comb_24h, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Comb_24h$S.Score
feature_counts_seurat_Comb_24h$G2M.Score

feature_counts_seurat_Comb_24h <- RunPCA(feature_counts_seurat_Comb_24h, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat_Comb_24h)

# Comb_48h 
intersect( g2m.genes, rownames(geneInfo_Comb_48h$external_gene_name))
intersect( s.genes, rownames(geneInfo_Comb_48h))

feature_counts_seurat_Comb_48h <- CellCycleScoring(feature_counts_seurat_Comb_48h, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Comb_48h$S.Score
feature_counts_seurat_Comb_48h$G2M.Score

feature_counts_seurat_Comb_48h <- RunPCA(feature_counts_seurat_Comb_48h, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat_Comb_48h)

#Ref
intersect( g2m.genes, rownames(geneInfo_Ref$external_gene_name))
intersect( s.genes, rownames(geneInfo_Ref))

feature_counts_seurat_Ref <- CellCycleScoring(feature_counts_seurat_Ref, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Ref@meta.data
#head(x = colnames(x = feature_counts_seurat_Ref))
feature_counts_seurat <- RunPCA(feature_counts_seurat_Ref, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat_Ref)

summary(feature_counts_seurat_Ref)

feature_counts_seurat_Ref@assays$RNA

# Dil 
intersect( g2m.genes, rownames(geneInfo_Dil$external_gene_name))
intersect( s.genes, rownames(geneInfo_Dil))

feature_counts_seurat_Dil <- CellCycleScoring(feature_counts_seurat_Dil, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Dil@meta.data
#head(x = colnames(x = feature_counts_seurat_Ref))
#feature_counts_seurat_Dil <- RunPCA(feature_counts_seurat_Dil, features = c(s.genes, g2m.genes))
#DimPlot(feature_counts_seurat_Dil)

summary(feature_counts_seurat_Dil)
feature_counts_seurat_Dil@assays$RNA

# Bulk 
intersect( g2m.genes, rownames(geneInfo_Bulk$external_gene_name))
intersect( s.genes, rownames(geneInfo_Bulk))

feature_counts_seurat_Bulk <- CellCycleScoring(feature_counts_seurat_Bulk, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat_Bulk@meta.data

summary(feature_counts_seurat_Bulk)

feature_counts_seurat_Bulk@assays$RNA
```


``` {r clusterData}
#RidgePlot(feature_counts_seurat_Comb, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
#s.genes
#intersect( g2m.genes, rownames(geneInfo$external_gene_name))

#intersect( s.genes, rownames(geneInfo))
#feature_counts_seurat <- CellCycleScoring(feature_counts_seurat, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
#feature_counts_seurat@meta.data
#RidgePlot(feature_counts_seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
#feature_counts_seurat <- RunPCA(feature_counts_seurat, features = c(s.genes, g2m.genes))
#DimPlot(feature_counts_seurat)
```

``` {r plotSeuratResult}
#Comb
VizDimLoadings(feature_counts_seurat_Comb, dims = 1:2, reduction = "pca")
DimPlot(feature_counts_seurat_Comb, reduction = "pca")
DimHeatmap(feature_counts_seurat_Comb, dims = 1:15, cells = 500, balanced = TRUE)
ElbowPlot(feature_counts_seurat_Comb)
feature_counts_seurat <- FindNeighbors(feature_counts_seurat_Comb, dims = 1:10)
#feature_counts_seurat <- FindClusters(feature_counts_seurat_Comb, resolution = 0.5)
head(Idents(feature_counts_seurat_Comb), 5)

#Ref
VizDimLoadings(feature_counts_seurat_Ref, dims = 1:2, reduction = "pca")
DimPlot(feature_counts_seurat_Ref, reduction = "pca")
DimHeatmap(feature_counts_seurat_Ref, dims = 1:15, cells = 500, balanced = TRUE)
ElbowPlot(feature_counts_seurat_Ref)
feature_counts_seurat <- FindNeighbors(feature_counts_seurat_Ref, dims = 1:10)
#feature_counts_seurat <- FindClusters(feature_counts_seurat_Comb, resolution = 0.5)
head(Idents(feature_counts_seurat_Ref), 5)
```

##Convert Seuratobject to dataframe

```{r Convert Seuratobject to dataframe}
#Convert Seuratobject to dataframe

#Comb
CombmetaData = feature_counts_seurat_Comb@meta.data
class(CombmetaData)
head(CombmetaData)

#Comb_0h
Comb_0h_metaData = feature_counts_seurat_Comb_0h@meta.data
class(Comb_0h_metaData)
head(Comb_0h_metaData)

#Comb_24h
Comb_24h_metaData = feature_counts_seurat_Comb_24h@meta.data
class(Comb_24h_metaData)
head(Comb_24h_metaData)

#Comb_48h
Comb_48h_metaData = feature_counts_seurat_Comb_48h@meta.data
class(Comb_48h_metaData)
head(Comb_48h_metaData)

#Ref
RefmetaData = feature_counts_seurat_Ref@meta.data
class(RefmetaData)
head(RefmetaData)

# Dil 
DilmetaData = feature_counts_seurat_Dil@meta.data
class(DilmetaData)
head(DilmetaData)

# Bulk 
BulkmetaData = feature_counts_seurat_Bulk@meta.data
class(BulkmetaData)
head(BulkmetaData)

#Count number of genes associated with each phase
table(feature_counts_seurat_Comb@meta.data$Phase)
table(feature_counts_seurat_Comb_0h@meta.data$Phase)
table(feature_counts_seurat_Comb_24h@meta.data$Phase)
table(feature_counts_seurat_Comb_48h@meta.data$Phase)
table(feature_counts_seurat_Ref@meta.data$Phase)
table(feature_counts_seurat_Dil@meta.data$Phase)
table(feature_counts_seurat_Bulk@meta.data$Phase)
#table(feature_counts_seurat@meta.data$Phase)

#Saving the result as a numeric vector
#Comb_numeric_vec <- as.vector(unlist(lapply(table(feature_counts_seurat_Comb@meta.data$Phase), as.numeric)))
#Ref_numeric_vec <- as.vector(unlist(lapply(table(feature_counts_seurat_Ref@meta.data$Phase), as.numeric)))
#Ref_numeric_vec
#Comb_numeric_vec
```
##Installing cyclone

``` {r load cyclone}
#Installing cyclone
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("scran")
#browseVignettes("scran")
```

``` {r installSingleCellExperiment}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("SingleCellExperiment")
```

## Check row names

``` {r checkRowNames}
rownames(geneInfo)
```

## Cyclone and preparation of data in order to run cyclone

``` {r Cyclone prep}
library(SingleCellExperiment)
library(scran)

?cyclone

#Comb
# Ändrar rownames till ensemblID för Comb
geneInfo_Comb$external_gene_name = rownames(geneInfo_Comb)
Comb = inner_join(geneInfo_Comb, geneInfoNames)
rownames(Comb) = Comb$Geneid
Comb2 = Comb %>% select(-Geneid,-external_gene_name)
dim(Comb)
dim(Comb2)
head(Comb2)

# Kontrollera att du har en rad med genenamn som första kolumn
Comb_new <- data.matrix(Comb2)
dim(Comb_new)

sce_Comb <- SingleCellExperiment (assays = list(counts = Comb_new, logcounts = log2(Comb_new+1)))

# Comb 0h 
# Ändrar rownames till ensemblID  för Comb_0h
geneInfo_Comb_0h$external_gene_name = rownames(geneInfo_Comb_0h)
Comb_0h = inner_join(geneInfo_Comb_0h, geneInfoNames)
rownames(Comb_0h) = Comb_0h$Geneid
Comb_0h2 = Comb_0h %>% select(-Geneid,-external_gene_name)
dim(Comb_0h)
dim(Comb_0h2)

# Kontrollera att du har en rad med genenamn som första kolumn
Comb_0h_new <- data.matrix(Comb_0h2)
dim(Comb_0h_new)

sce_Comb_0h <- SingleCellExperiment (assays = list(counts = Comb_0h_new, logcounts = log2(Comb_0h_new+1)))

# Comb 24h 
# Ändrar rownames till ensemblID för Comb_24h
geneInfo_Comb_24h$external_gene_name = rownames(geneInfo_Comb_24h)
Comb_24h = inner_join(geneInfo_Comb_24h, geneInfoNames)
rownames(Comb_24h) = Comb_24h$Geneid
Comb_24h2 = Comb_24h %>% select(-Geneid,-external_gene_name)
dim(Comb_24h)
dim(Comb_24h2)

# Kontrollera att du har en rad med genenamn som första kolumn
Comb_24h_new <- data.matrix(Comb_24h2)
dim(Comb_24h_new)

sce_Comb_24h <- SingleCellExperiment (assays = list(counts = Comb_24h_new, logcounts = log2(Comb_24h_new+1)))

#Comb 48h
# Ändrar rownames till ensemblID för Comb_48h
geneInfo_Comb_48h$external_gene_name = rownames(geneInfo_Comb_48h)
Comb_48h = inner_join(geneInfo_Comb_48h, geneInfoNames)
rownames(Comb_48h) = Comb_48h$Geneid
Comb_48h2 = Comb_48h %>% select(-Geneid,-external_gene_name)
dim(Comb_48h)
dim(Comb_48h2)

# Kontrollera att du har en rad med genenamn som första kolumn
Comb_48h_new <- data.matrix(Comb_48h2)
dim(Comb_48h_new)

sce_Comb_48h <- SingleCellExperiment (assays = list(counts = Comb_48h_new, logcounts = log2(Comb_48h_new+1)))

# Ref
#Ändrar rownames till ensemblID för Ref
geneInfo_Ref$external_gene_name = rownames(geneInfo_Ref)
Ref = inner_join(geneInfo_Ref, geneInfoNames)
rownames(Ref) = Ref$Geneid
Ref2 = Ref %>% select(-Geneid,-external_gene_name)
dim(Ref)
dim(Ref2)

# Kontrollera att du har en rad med genenamn som första kolumn
Ref_new <- data.matrix(Ref2)
dim(Ref_new)

sce_Ref <- SingleCellExperiment (assays = list(counts = Ref_new, logcounts = log2(Ref_new+1)))

```

## Run Cyclone

``` {r Run Cyclone}

mmu.pairs <- readRDS(system.file("exdata", "human_cycle_markers.rds", package="scran"))

# Comb
sce_Comb <- SingleCellExperiment (assays = list(counts = Comb_new, logcounts = log2(Comb_new+1)))
cc_Comb<- cyclone(sce_Comb, pairs=mmu.pairs, gene.names = rownames(Comb_new))
table(cc_Comb$phase)

# Comb_0h
sce_Comb_0h <- SingleCellExperiment (assays = list(counts = Comb_0h_new, logcounts = log2(Comb_0h_new+1)))
cc_Comb_0h<- cyclone(sce_Comb_0h, pairs=mmu.pairs, gene.names = rownames(Comb_0h_new))
table(cc_Comb_0h$phase)

# Comb_24h
sce_Comb_24h <- SingleCellExperiment (assays = list(counts = Comb_24h_new, logcounts = log2(Comb_24h_new+1)))
cc_Comb_24h<- cyclone(sce_Comb_24h, pairs=mmu.pairs, gene.names = rownames(Comb_24h_new))
table(cc_Comb_24h$phase)

# Comb_48h
sce_Comb_48h <- SingleCellExperiment (assays = list(counts = Comb_48h_new, logcounts = log2(Comb_48h_new+1)))
cc_Comb_48h<- cyclone(sce_Comb_48h, pairs=mmu.pairs, gene.names = rownames(Comb_48h_new))
table(cc_Comb_48h$phase)

# Ref
sce_Ref <- SingleCellExperiment (assays = list(counts = Ref_new, logcounts = log2(Ref_new+1)))
cc_Ref<- cyclone(sce_Ref, pairs=mmu.pairs, gene.names = rownames(Ref_new))
table(cc_Ref$phase)

``` 
## Plotta Cycloneresultat

``` {r Plot Cyclone}

# Comb
cc_Comb
class(cc_Comb)
#table(cc_Comb$phase)

# Comb_0h
cc_Comb_0h
#table(cc_Comb_0h$phase)

# Comb_24h
cc_Comb_24h
#table(cc_Comb_24h$phase)

# Comb_48h
cc_Comb_48h
#table(cc_Comb_48h$phase)

# Ref
cc_Ref
#table(cc_Ref$phase)

``` 
## Korrigera pairs

``` {r Pairs}
# Iterera över alla gener och gör en jämförelse

#mmu.pairs.reduced  = mmu.pairs
#mmu.pairs.reduced$G1 = mmu.pairs.reduced$G1[mmu.pairs$G1$first == "ENSG00000100519",]


#findBestGene <- function(cc_ref, cc_comb_0h,mmu.pairs, alreadyDecidedGenes){
  
  # Compare distribution to find most similair
    
  # Compare diluted and facs to see if scores are the same. (should be the same for 0h_ref,0h_comb, 24h_comb, 48h_cpomb) 
  # Make some choice on what gene to choose
  
 # decidedGenes = c(alreadyDecidedGenes, newGene)
#  return ()
#} 
``` 
