---
title: "Seurat_protocol_2"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library

```{r library}
library(dplyr)
library(Seurat)
library(data.table)
# load data from Seurat package
#data(cc.genes)

# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

## Load the dataset

```{r load}
feature_counts.data <- read.table(file = "/Users/linneaeriksson/Desktop/Tillämpad bioinformatik/scrna-cell-cycle/feature_counts.table.tsv", 
                                  sep = "\t",
                                  header = TRUE,
                                  quote = "",
                                  stringsAsFactors = FALSE)
dim(x=feature_counts.data)
head(x = colnames(x = feature_counts.data))
```

## Remove data that is not sc from dataset

```{r sc}
removeIDs <- function(feature_counts.data) {
  column_indices_ID <- c()
  
  for(i in 1:ncol(feature_counts.data)) {
    if(grepl(".dil.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
    if(grepl(".FACS.", colnames(feature_counts.data[i]))) {
      column_indices_ID <- c(column_indices_ID, i)
    }
  }
  feature_counts_removedColumns <- c()
  feature_counts_removedColumns <- feature_counts.data[,-column_indices_ID]
  return(feature_counts_removedColumns)
}

feature_counts_SC <- removeIDs(feature_counts.data)
dim(feature_counts_SC)
head(x = colnames(x = feature_counts_SC))

```

## Remove columns that are not of interest from dataset

```{r removeColumns}
feature_counts_SC= feature_counts_SC[,-2:-6]
dim(feature_counts_SC)
head(x = colnames(x = feature_counts_SC))
```

## Load file for correct gene names

```{r load2}
geneInfoFile <- read.table(file = "/Users/linneaeriksson/Desktop/Tillämpad bioinformatik/scrna-cell-cycle/geneInfo.QC.method.seurat.SCORPIUS.tsv", 
                           sep = "\t",
                           header = TRUE,
                           quote = "",
                           stringsAsFactors = FALSE)

#only protein coding genes
geneInfoFile = geneInfoFile[geneInfoFile$gene_biotype == "protein_coding",] 

#only genes represented with one unique ensembl ID

tmp = data.frame(table(geneInfoFile$external_gene_name))
tmp = tmp[tmp$Freq ==1, ]
geneInfoFile = geneInfoFile[geneInfoFile$external_gene_name %in% tmp$Var1, ]


dim(x=geneInfoFile)

head(geneInfoFile)
```

## Select relevant columns

```{r select}

geneInfoNames = geneInfoFile %>%  select(Geneid,external_gene_name)

head(geneInfoNames)
dim(x=geneInfoNames)

intersect( g2m.genes, geneInfoNames$external_gene_name)
#intersect( s.genes, rownames(geneInfo))


## Remove Chr with NA
```

``` {r NA}
#geneInfoNames = geneInfoNames[-52747:-58303,]
#head(geneInfoNames)
#dim(geneInfoNames)


```


## Perform inner join 

```{r innerJoin}
geneInfo = inner_join(geneInfoNames, feature_counts_SC)

intersect( g2m.genes, geneInfo$external_gene_name)


dim(x=geneInfo)
```

## Set rownames to correct gene names

```{r rownames}

rownames(geneInfo) =  make.names(geneInfo$external_gene_name, unique =TRUE)

#test = data.frame(make.names(geneInfo$Chr, unique =TRUE)
#length(test)

rownames(geneInfo) =  geneInfo$external_gene_name

dim(x=geneInfo)
head(x = colnames(x = geneInfo))

```

## Remove columns in order to perform rowSums

```{r removeColumns2}
geneInfo= geneInfo[,-1:-2]
dim(geneInfo)
head(x = colnames(x = geneInfo))

intersect( g2m.genes, rownames(geneInfo))

```

## Remove genes that are not expressed

```{r expressed}

#rowSums(geneInfo)
keep <- rowSums(geneInfo)>1
geneInfo <- geneInfo[keep,]
dim(geneInfo)

intersect( g2m.genes, rownames(geneInfo))


```

## Create a Seurat object out of the table

```{r Seurat}
feature_counts_seurat <- CreateSeuratObject(counts = geneInfo, min.cells = 3)
feature_counts_seurat
feature_counts_seurat <- NormalizeData(object = feature_counts_seurat)
feature_counts_seurat <- FindVariableFeatures(object = feature_counts_seurat, selection.method = "vst")
feature_counts_seurat <- ScaleData(feature_counts_seurat, features = rownames(feature_counts_seurat))


feature_counts_seurat <- RunPCA(object = feature_counts_seurat, pc.genes = feature_counts_seurat@var.genes, do.print = FALSE)
feature_counts_seurat
#DimHeatmap(feature_counts_seurat)

DimHeatmap(feature_counts_seurat)

```

## Assign cell-cycle scores

```{r Scoring}

feature_counts_seurat <- CellCycleScoring(object = feature_counts_seurat, s.features = s.genes, g2m.features = g2m.genes)


#class(geneInfoFile)

#geneInfoFile[3,5]

#geneInfo["CFTR",]


feature_counts_seurat@meta.data

table(feature_counts_seurat@meta.data$Phase)
head(feature_counts_seurat[[]])
feature_counts_seurat@meta.data$Phase
```


``` {r clusterData}



RidgePlot(feature_counts_seurat, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)


s.genes

intersect( g2m.genes, rownames(geneInfo$external_gene_name))
intersect( s.genes, rownames(geneInfo))

feature_counts_seurat <- CellCycleScoring(feature_counts_seurat, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

feature_counts_seurat@meta.data
feature_counts_seurat <- RunPCA(feature_counts_seurat, features = c(s.genes, g2m.genes))
DimPlot(feature_counts_seurat)

# Save seurat object as file
#saveRDS(feature_counts_seurat,file="/Users/annaasklof/Desktop/Seurat/S/feature_counts_seurat.rds")
resultSeurat <- as.data.frame(as.matrix(feature_counts_seurat@meta.data))
fwrite(x = resultSeurat, row.names = TRUE, col.names =TRUE, file = "/Users/annaasklof/Desktop/Seurat/S/resultSeurat.csv")
```

## Assign cell-cycle scores

```{r Comparing with the ground truth data}
correctCCS <- read.table(file = "/Users/annaasklof/Desktop/Seurat/S/Data/sampleData.QC.method.tsv", 
                                  sep = "\t",
                                  header = TRUE,
                                  quote = "",
                                  stringsAsFactors = FALSE)
dim(x=correctCCS)
head(x = colnames(x = correctCCS))

#select relevant colums in the correctCCS file
correctCCS = correctCCS %>%  select(fileName, CellCycle)

dim(x=correctCCS)
head(x = colnames(x = correctCCS))

# Select relevant colums from feature_counts_seurat
result_feature_counts <- read.table(file = "/Users/annaasklof/Desktop/Seurat/S/resultSeurat.csv", 
                                  sep = "\t",
                                  header = TRUE,
                                  quote = "",
                                  stringsAsFactors = FALSE)
dim(x=result_feature_counts)
head(x = colnames(x = result_feature_counts))
result_feature_counts

#select relevant colums from the result
#result_feature_counts = result_feature_counts %>%  select(X, Phase)

# Compare the cell cycle state predicted to the ground throuth
#CellCycleStates = inner_join(sampleData, result_feature_counts) 


```


``` {r plotSeuratResult}
VizDimLoadings(feature_counts_seurat, dims = 1:2, reduction = "pca")
DimPlot(feature_counts_seurat, reduction = "pca")
DimHeatmap(feature_counts_seurat, dims = 1:15, cells = 500, balanced = TRUE)

ElbowPlot(feature_counts_seurat)

feature_counts_seurat <- FindNeighbors(feature_counts_seurat, dims = 1:10)
feature_counts_seurat <- FindClusters(feature_counts_seurat, resolution = 0.5)
head(Idents(feature_counts_seurat), 5)

```

